### 2021-07-13 chapters/1_abstract.tex
Fixing typos styling etc..

\documentclass[../main]{subfiles}
\begin{document}
\chapter*{Abstract}
Software testing is a crucial part of the software engineering process. 
It is widely used in the industry.
A part of software testing is building test suites which contain unit tests.
These unit tests are written by developers. 
As projects grow the test suite grows along.
Maintaining and monitoring these test suites is important because they influence the cost of maintenance.
Monitoring the quality of test code has shown that tests with lower quality leads to more defect-prone production code.
Measuring the effectiveness of test suites detecting faults is part of the monitoring process.
A technique to measure the test effectiveness is mutation testing.
Mutation testing is a computationally expensive technique that requires the ability to run tests.
Different techniques have been proposed to reduce the cost of mutation testing.
One of these techniques is the clustering of mutants.
By clustering the mutants we can execute less mutants to reduce the cost.
We conducted two experiment in which we cluster the mutants with two different techniques.
Before conducting the experiment we did a state of the art comparison between three mutation testing tools for Java.
Based on this research we selected a mutation testing tool to use in our experiments.
\newline
We started our experiment by identifying characteristics to represent the mutants.
For the first experiment we used hierarchical clustering and in the second experiment we trained the fuzzy c-means model.
We calculated a weighted mutation score and compared this with the mutation score of a full set of mutants executed.
Results show that with hierarchical clustering we can reduce the amount of mutants executed while maintaining the effectiveness.
The clusters generated by the machine learning model were less accurate and showed a significant effectiveness decrease.
The machine learning model didn't perform as well as the hierarchical clustering.
However, more work is needed in finding the right parameters for the fuzzy c-means algorithm. 
\end{document}

>>>>

\documentclass[../main]{subfiles}
\begin{document}
\chapter*{Abstract}
Software testing is a crucial part of the software engineering process. 
It is widely used in the industry.
A part of software testing is building test suites which contain unit tests.
These unit tests are written by developers. 
As projects grow the test suite grows along.
Maintaining and monitoring these test suites is important because they influence the cost of maintenance.
Monitoring the quality of test code has shown that tests with lower quality leads to more defect-prone production code.
\newline
Measuring the effectiveness of test suites detecting faults is part of the monitoring process.
A technique to measure the test effectiveness is mutation testing.
Mutation testing is a computationally expensive technique that requires the ability to run tests.
Different techniques have been proposed to reduce the cost of mutation testing.
One of these techniques is the clustering of mutants.
By clustering mutants we can execute less mutants to reduce the cost.
\newline
We conducted two experiment in which we cluster the mutants with two different techniques.
Before conducting the experiment we conducted a state of the art comparison between three mutation testing tools for Java.
Based on this research we selected a mutation testing tool to use in our experiments.
\newline
We started our experiment by identifying characteristics to represent the mutants.
For the first experiment we used hierarchical clustering and in the second experiment we trained the fuzzy c-means model.
We calculated a weighted mutation score and compared this with the mutation score of a full set of mutants executed.
Results show that with hierarchical clustering we can reduce the amount of mutants executed while maintaining the effectiveness.
The clusters generated by the machine learning model were less accurate and showed a significant decrease in effectiveness.
The machine learning model didn't perform as well as the hierarchical clustering.
However, more work is needed in finding the right parameters for the fuzzy c-means algorithm. 
\end{document}

### 2021-07-13 chapters/2_introduction.tex
Fixing typos styling etc..

\chapter{Introduction}
\label{ch:introduction}
Software testing is a critical part in the software engineering process. It is widely used and accepted in the industry for quality assurance. 
Software testing can detect software bugs during the development process and can also serve for regression purposes\cite{Bertolino2007SoftwareDreams}. 
During the software testing process developers write automated tests called unit tests. 
These tests can be run by testing frameworks such as JUnit and TestNG\cite{JUnit, TestNG}.
Monitoring the quality of test code has shown that tests with lower quality leads to more defect-prone production code\cite{Spadini2018OnQuality}. 
Research has also found that if the production code grows, the test code grows along with it\cite{Zaidman2008MiningCode}. 
Quality control on test suites is therefore important as the maintenance on tests can be difficult and generate risks if done incorrectly\cite{Horvath2015TestMetrics}.
\newline
The growing complexity and size of code bases consequently lead to incomprehensible tests.
As a result test bugs can occur. 
A test bug is a test that fail even tough the program is correct(\textit{false positive}). 
Another bug, which is even worse, are tests that do not fail when the program is not working as desired(\textit{false negative}).
The \textit{false negative} is a problem when breaking changes or not detected by the test suite.
\newline
To solve this issue we can measure the fault detecting capability of a test suite, this is called the test suite effectiveness.
Test suite effectiveness is measured by the number of faulty versions of a \acrfull{sut} that are detected by a test suite.
Real faults are unknown in advance, we can use mutation testing as a proxy measurement. 
Research has found statistically significant evidence which correlates mutant detection with real fault detection\cite{Just2014AreTesting}.
\newline
Mutation testing tools create small faulty versions of the program.
The test suite is then run to verify if these faults are detected.
These faults are called mutants and are created by mutators.
Mutators mutate specific expressions or statements in the source code. 
To prevent changing the overall functionality of the program, mutants are represented as very small changes\cite{Lipton1978HintsProgrammer}.
\newline
After all the mutants are executed the mutation score can be calculated. 
The mutation score is the percentage of mutants killed divided by the total amount of mutants executed.
If a large amount of mutants survives, it might be an indication that the quality of the test suite is poor as programming errors remain undetected.
Mutation testing has considerable drawbacks, such as the equivalent mutant problem and being resource expensive.

\section{Problem statement}
Mutation testing requires a significant amount of computational resources\cite{Kintis2016EffectiveMutation}. 
The high cost requirement is often a barrier for adopting mutation testing \cite{Pizzoleto2019}.
A lot of techniques and methods have been developed to improve the performance, however, most of these approaches are not as effective as mutation testing a full set of mutants\cite{Pizzoleto2019,Yao2014}. 
\newline
A technique to reduce the number of mutants or the number of test case executions is mutation clustering.
Mutation clustering aims to reduce the amount of mutants to be executed by clustering the mutants\cite{Ma2016,Yu2019PossibilityScope}.
Clustering is a technique used for the grouping of data, for pattern recognition and classification, based on some similar characteristics\cite{Alsabti1997}.
The clustering of mutants has been researched with promising results\cite{Ji2009,Wilinski2015,Ma2016}. For example Ma et al., \cite{Ma2016} clustered the mutants for expressions. 
However in their paper they indicate that the choosing expressions as the domain of their cluster is a limitation.
A recurring problem in the research of mutation clustering is to determine the means of domains as the centre of a cluster\cite{Ji2009,Wilinski2015,Ma2016,Wei2021SpectralTesting}.
\newline
To reduce the cost of mutation testing, we try to find a solution that can cluster mutants while maintaining the accuracy of the same complete set of mutants. 
Our goal is to cluster any mutant that is generated.
We will do this with two different approaches.
Our first approach will be a qualitative approach and second a quantitative approach. 
We structure our research on the following research questions:
\newline
\textbf{Research Question 1}: What set of characteristics can we identify for clustering generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
\textbf{Research Question 1.1}: How do the existing mutation testing tools for Java compare to each other?
\newline
\textbf{Research Question 2}: How can we train a machine learning model to recognize and cluster generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
We start with comparing existing mutation testing tools. 
We will use the existing literature as a starting point and research what would be the best tool to use for this thesis.
\newline
After we have selected a tool we will research existing clustering techniques.
Based on the results of this research we select and implement a clustering algorithm that will cluster a characteristic of the mutants. 
We will randomly select on mutant per of each cluster to be executed by the mutation tooling.
We will then compare the mutation score of the clustered mutants with the mutation score of a full set of mutants.
\newline
To answer our second research question we will train a machine learning model to cluster mutants.
We can then repeat the comparison we did for research question one. 
\newline
We want to see how the algorithm/model performs on big and small projects.
These project must also require a test suite.
\newline
We chose to focus on projects that use Java as main programming language because it is one of the most popular languages\cite{Tiobe}.

\section{Contributions}
Our research makes the following contributions:
\begin{enumerate}
 \item A methodology to cluster generated mutants based on chosen characteristics to reduce the cost of mutation testing.
 \item A \acrfull{poc} which implements the qualitative methodology chosen and elaborated within the thesis.
 \item A \acrfull{poc} which implements the quantitative approach to cluster mutants. 
\end{enumerate}

\section{Outline}
In Chapter~\ref{ch:background} we describe the background of this thesis. 
Chapter \ref{ch:mutation-tool} compares the different mutation testing tools to decide on a tool to use for this research.
In chapter \ref{ch:reserach_question1} we identify characteristics to represent mutants and design an experiment to cluster them.
In chapter \ref{ch:reserach_question2} we select a machine learning model and design an experiment to cluster the mutants with this model.
Results are shown in Chapters \ref{ch:results_rq1} and \ref{ch:results_rq2} for the experiments in described Chapters \ref{ch:reserach_question1} and \ref{ch:reserach_question2} respectively. 
The results are then discussed in Chapter~\ref{ch:discussion}. Chapter~\ref{ch:related_work}, contains the work related to this thesis.
Finally, we present our concluding remarks in Chapter~\ref{ch:conclusion} together with future work.

>>>>

\chapter{Introduction}
\label{ch:introduction}
Software testing is a critical part in the software engineering process. It is widely used and accepted in the industry for quality assurance. 
Software testing can detect software bugs during the development process and can also serve for regression purposes\cite{Bertolino2007SoftwareDreams}. 
During the software testing process developers write automated tests called unit tests. 
These tests can be run by testing frameworks such as JUnit and TestNG\cite{JUnit, TestNG}.
Monitoring the quality of test code has shown that tests with lower quality leads to more defect-prone production code\cite{Spadini2018OnQuality}. 
Research has also found that if the production code grows, the test code grows along with it\cite{Zaidman2008MiningCode}. 
Quality control on test suites is therefore important as the maintenance on tests can be difficult and generates risks if done incorrectly\cite{Horvath2015TestMetrics}.
\newline
The growing complexity and size of code bases consequently lead to incomprehensible tests.
As a result test bugs can occur. 
A test bug is a test that fail even tough the program is correct(\textit{false positive}). 
Another bug, which is even worse, are tests that do not fail when the program is not working as desired(\textit{false negative}).
The \textit{false negative} is a problem when breaking changes are not detected by the test suite.
\newline
To solve this issue we can measure the fault detecting capability of a test suite, this is called the test suite effectiveness.
Test suite effectiveness is measured by the number of faulty versions of a \acrfull{sut} that are detected by a test suite.
Real faults are unknown in advance, we can use mutation testing as a proxy measurement. 
Research has found statistically significant evidence which correlates mutant detection with real fault detection\cite{Just2014AreTesting}.
\newline
Mutation testing tools create small faulty versions of the program.
The test suite is then run to verify if these faults are detected.
These faults are called mutants and are created by mutators.
Mutators mutate specific expressions or statements in the source code. 
To prevent changing the overall functionality of the program, mutants are represented as very small changes\cite{Lipton1978HintsProgrammer}.
\newline
After all the mutants are executed the mutation score can be calculated. 
The mutation score is the percentage of mutants killed divided by the total amount of mutants executed.
If a large amount of mutants survives, it might be an indication that the quality of the test suite is poor as programming errors remain undetected.
Mutation testing has considerable drawbacks, such as the equivalent mutant problem and being resource expensive.

\section{Problem statement}
Mutation testing requires a significant amount of computational resources\cite{Kintis2016EffectiveMutation}. 
The high cost requirement is often a barrier for adopting mutation testing \cite{Pizzoleto2019}.
A lot of techniques and methods have been developed to improve the performance, however, most of these approaches are not as effective as mutation testing a full set of mutants\cite{Pizzoleto2019,Yao2014}. 
\newline
A technique to reduce the number of mutants or the number of test case executions is mutation clustering.
Mutation clustering aims to reduce the amount of mutants to be executed by clustering the mutants\cite{Ma2016,Yu2019PossibilityScope}.
Clustering is a technique used for the grouping of data, for pattern recognition and classification, based on some similar characteristics\cite{Alsabti1997}.
The clustering of mutants has been researched with promising results\cite{Ji2009,Wilinski2015,Ma2016}. For example Ma et al., \cite{Ma2016} clustered the mutants for expressions. 
However in their paper they indicate that the choosing expressions as the domain of their cluster is a limitation.
A recurring problem in the research of mutation clustering is to determine the means of domains as the centre of a cluster\cite{Ji2009,Wilinski2015,Ma2016,Wei2021SpectralTesting}.
\newline
To reduce the cost of mutation testing, we try to find a solution that can cluster mutants while maintaining the accuracy of the same complete set of mutants. 
Our goal is to cluster any mutant that is generated.
We will do this with two different approaches.
Our first approach will be a qualitative approach and second a quantitative approach. 
We structure our research on the following research questions:
\newline
\textbf{Research Question 1}: What set of characteristics can we identify for clustering generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
\textbf{Research Question 1.1}: How do the existing mutation testing tools for Java compare to each other?
\newline
\textbf{Research Question 2}: How can we train a machine learning model to recognize and cluster generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
We start with comparing existing mutation testing tools. 
We will use the existing literature as a starting point and research what would be the best tool to use for this thesis.
\newline
After we have selected a tool we will research existing clustering techniques.
Based on the results of this research we select and implement a clustering algorithm that will cluster one ore more characteristics of the mutants. 
We will randomly select on mutant per of each cluster to be executed by the mutation tooling.
We will then compare the mutation score of the clustered mutants with the mutation score of a full set of mutants.
\newline
To answer our second research question we will train a machine learning model to cluster mutants.
We can then repeat the comparison we did for research question one. 
\newline
We want to see how the algorithm/model performs on big and small projects.
These project must also require a test suite.
\newline
We chose to focus on projects that use Java as main programming language because it is one of the most popular languages\cite{Tiobe}.

\section{Contributions}
Our research makes the following contributions:
\begin{enumerate}
 \item A methodology to cluster generated mutants based on chosen characteristics to reduce the cost of mutation testing.
 \item A \acrfull{poc} which implements the qualitative methodology chosen and elaborated within the thesis.
 \item A \acrfull{poc} which implements the quantitative approach to cluster mutants and is elaborated within the thesis. 
\end{enumerate}

\section{Outline}
In Chapter~\ref{ch:background} we describe the background of this thesis. 
Chapter \ref{ch:mutation-tool} compares the different mutation testing tools to decide on a tool to use for this research.
In chapter \ref{ch:reserach_question1} we identify characteristics to represent mutants and design an experiment to cluster them.
In chapter \ref{ch:reserach_question2} we select a machine learning model and design an experiment to cluster the mutants with this model.
Results are shown in Chapters \ref{ch:results_rq1} and \ref{ch:results_rq2} for the experiments described in Chapters \ref{ch:reserach_question1} and \ref{ch:reserach_question2} respectively. 
The results are then discussed in Chapter~\ref{ch:discussion}.
\newline
Chapter~\ref{ch:related_work}, contains the work related to this thesis.
Finally, we present our concluding remarks in Chapter~\ref{ch:conclusion} together with future work.

### 2021-07-13 chapters/3_background.tex
fixed our tool to the tool
Fixing typos styling etc..

\chapter{Background}
\label{ch:background}
This chapter will present the necessary background information for this thesis. First, we define some basic terminology that will be used throughout this thesis.
Next, we will compare existing mutation testing tools. We start of by reviewing existing literature and then extend the comparison. \todo{extend this later on?}
\section{Terminology}
\todo{do this when we finished experiments and all}
\todo{subsuming mutants}
\todo{levensteihn distance}
\todo{maven and gradle}
\section{Project selection}
\label{ch:project_selection}
We choose three main requirements for selecting software projects; the projects should have a test suite, the test suite should not contain failing tests and the mutation testing tool should be able to execute mutants for the sample project.
We selected projects that were also used in other research within the mutation testing and testing domain\cite{Pizzoleto2019,Yu2019PossibilityScope,Wei2021SpectralTesting, Zhang2019PredictiveTesting, Chen2018SpeedingStudy, Laurent2017AssessingPIT}.
We extend our sample by selecting projects from the first six pages of the most popular Java projects on GitHub\footnote{\url{https://github.com/search?p=7&q=language\%3Ajava+stars\%3A\%3E10000&type=Repositories}}.
The unfiltered sample contains 50+ projects.
From there we filter out all the projects that contain failing tests and that aren't libraries or applications.
The result consists of a sample with fifteen projects in total.

\section{Clustering algorithms}
This research focuses on clustering mutants and use of existing clustering algorithms. 
Many different types of clustering algorithms have been proposed\cite{Rodriguez2019}. While there is a lot of diversity, some methods are more frequently used than others\cite{Wu2008TopMining}. 
Rodriguez et al., compared the performance of nine different clustering algorithms:
\begin{itemize}
    \item k-means
    \item CLARA
    \item hierarchical
    \item EM
    \item hcmodel
    \item spectral
    \item subscpace
    \item optics
    \item dbscan
\end{itemize}
The goal of their study was to guide researchers, who have little experience in data mining techniques, to the application of clustering methods.
They evaluated the algorithms in three distinct situations: default parameters, single parameter variation and random variation of parameters.
They used 400 generated artificial data sets which were normally distributed.
\newline
The results reported in their research are respective to specific configurations of normally distributed data and algorithmic implementations.
Nonetheless they do give a good overview on how the algorithms compare to each other.
\newline
For the default parameter situation the spectral clustering algorithm had the best performance and the hierarchical algorithm had the worst performance.
\newline
Regarding single parameter variations, for data sets containing 2 columns, the hierarchical, optics and EM methods showed significant performance variation.
\newline
With respect to the multidimensional analysis for data sets, the performance of the algorithms for the multidimensional selection of parameters was similar to that using the default parameters.
They conclude their research with observing that, for data sets with 10 or more columns the spectral algorithm consistently provided the best performance.
However the EM, hierarchical, k- means and subspace algorithms can also achieve similar performance with some parameter tuning.
The optics and dbscan algorithms aim at different data distributions than performed in this study. There different results could be obtained for non-normally distributed data.

\section{Machine learning models for clustering}
\label{ch:neuralNetworkSurvey}
There are many different types of machine learning models with their respective applications.
For this thesis we will focus on machine learning models designed for clustering.
\newline
Machine learning algorithms are divided in supervised and unsupervised learning algorithms\cite{supervisedUnsupervised}.
Supervised learning is the machine learning task of learning a function that maps an input to an output based on example input-output pairs\cite{supervisedUnsupervised}.
In contrast to supervised learning, unsupervised learning shows self-organization that captures patterns as neuronal predilections or probability densities\cite{supervisedUnsupervised}.
Unsupervised learning algorithms are commonly used for classification and categorization\cite{supervisedUnsupervised}.
Unsupervised learning is a type of algorithm that learns patterns from untagged data\cite{supervisedUnsupervised}.
These machine learning algorithms are commonly used for clustering and dimensionality reduction\cite{supervisedUnsupervised}.
\newline
K.L. Du surveyed different machine learning models used for clustering\cite{Du2010Clustering:Approach}.
The models he discusses are:
\begin{itemize}
    \item Self organizing map.
    \item Learning vector quantization.
    \item C-means(K-means) clustering.
    \item Mountain and subtractive clustering.
    \item Neural gas.
    \item ART networks.
    \item Fuzzy c-means.
\end{itemize}
K.L. Du describes the mathematical principles on which the machine learning algorithms are based as well as the origin, pros and cons of each model.
\newline
He also touches on subjects relevant to clustering and machine learning models such as the under-utilization problem, fuzzy clustering, robust clustering, clustering based on non- Euclidean distance measures, supervised clustering, and hierarchical clustering.
Machine learning model variants and their references are also discussed.
\newline
He closes his paper with computer simulations of some of the machine learning models he surveyed.

>>>>

\chapter{Background}
\label{ch:background}
This chapter will present the necessary background information for this thesis. First, we define some basic terminology that will be used throughout this thesis.
Next, we will compare existing mutation testing tools. We start of by reviewing existing literature and then extend the comparison. 
This thesis relies on clustering techniques.
We review the state of the art on clustering algorithms and secondly the state of the art on machine learning models specialized in clustering.

\section{Terminology}
Subsuming mutants.
Mutation testing.
Levenshtein distance.
Maven.
Gradle.
Mutator.
test engine??
\section{Project selection}
\label{ch:project_selection}
We choose three main requirements for selecting software projects; the projects should have a test suite, the test suite should not contain failing tests and the mutation testing tool should be able to execute mutants for the sample project.
We selected projects that were also used in other research within the mutation testing and testing domain\cite{Pizzoleto2019,Yu2019PossibilityScope,Wei2021SpectralTesting, Zhang2019PredictiveTesting, Chen2018SpeedingStudy, Laurent2017AssessingPIT}.
\newline
We extend our sample by selecting projects from the first six pages of the most popular Java projects on GitHub\footnote{\url{https://github.com/search?p=7&q=language\%3Ajava+stars\%3A\%3E10000&type=Repositories}}.
The unfiltered sample contains 50+ projects.
From there we filter out all the projects that contain failing tests and that aren't libraries or applications.
The result consists of a sample with fifteen projects in total.

\section{Clustering algorithms}
This research focuses on clustering mutants and use of existing clustering algorithms. 
Many different types of clustering algorithms have been proposed\cite{Rodriguez2019}. While there is a lot of diversity, some methods are more frequently used than others\cite{Wu2008TopMining}. 
Rodriguez et al., compared the performance of nine different clustering algorithms:
\begin{itemize}
    \item k-means
    \item CLARA
    \item hierarchical
    \item EM
    \item hcmodel
    \item spectral
    \item subscpace
    \item optics
    \item dbscan
\end{itemize}
The goal of their study was to guide researchers, who have little experience in data mining techniques, to the application of clustering methods.
They evaluated the algorithms in three distinct situations: default parameters, single parameter variation and random variation of parameters.
They used 400 generated artificial data sets which were normally distributed.
\newline
The results reported in their research are respective to specific configurations of normally distributed data and algorithmic implementations.
Nonetheless they do give a good overview on how the algorithms compare to each other.
\newline
For the default parameter experiment, the spectral clustering algorithm had the best performance and the hierarchical algorithm had the worst performance.
\newline
Regarding single parameter variations, for data sets containing 2 columns, the hierarchical, optics and EM methods showed significant performance variation.
\newline
With respect to the multidimensional analysis for data sets, the performance of the algorithms for the multidimensional selection of parameters was similar to that using the default parameters.
They conclude their research with observing that, for data sets with 10 or more columns the spectral algorithm consistently provided the best performance.
However the EM, hierarchical, k- means and subspace algorithms can also achieve similar performance with some parameter tuning.
The optics and dbscan algorithms aim at different data distributions than performed in this study. There different results could be obtained for non-normally distributed data.

\section{Machine learning models for clustering}
\label{ch:neuralNetworkSurvey}
There are many different types of machine learning models with their respective applications.
For this thesis we will focus on machine learning models designed for clustering.
\newline
Machine learning algorithms are divided in supervised and unsupervised learning algorithms\cite{supervisedUnsupervised}.
Supervised learning is the machine learning task of learning a function that maps an input to an output based on example input-output pairs\cite{supervisedUnsupervised}.
In contrast to supervised learning, unsupervised learning shows self-organization that captures patterns as neuronal predilections or probability densities\cite{supervisedUnsupervised}.
Unsupervised learning algorithms are commonly used for classification and categorization\cite{supervisedUnsupervised}.
Unsupervised learning is a type of algorithm that learns patterns from untagged data\cite{supervisedUnsupervised}.
These machine learning algorithms are commonly used for clustering and dimensionality reduction\cite{supervisedUnsupervised}.
\newline
K.L. Du surveyed different machine learning models used for clustering\cite{Du2010Clustering:Approach}.
Relevant models he discusses are:
\begin{itemize}
    \item Self organizing map.
    \item Learning vector quantization.
    \item C-means(K-means) clustering.
    \item Mountain and subtractive clustering.
    \item Neural gas.
    \item ART networks.
    \item Fuzzy c-means.
\end{itemize}
K.L. Du describes the mathematical principles on which the machine learning algorithms are based as well as the origin, pros and cons of each model.
\newline
He also touches on subjects relevant to clustering and machine learning models such as the under-utilization problem, fuzzy clustering, robust clustering, clustering based on non- Euclidean distance measures, supervised clustering, and hierarchical clustering.
Machine learning model variants and their references are also discussed.
\newline
He closes his paper with computer simulations of some of the machine learning models he surveyed.

### 2021-07-13 chapters/4_mutation_tool.tex
changes
Fixing typos styling etc..

\chapter{Mutation testing tools}
\label{ch:mutation∂-tool}

Before we can start clustering mutants we need to select a tool that helps us generate mutants. 
Our goal is to find the tool that generates the greatest amount of mutants. 
\newline
First, we review existing literature. 
Secondly we extend the comparison by adding information about the current state of the tools.
With the combination of the information gathered we make a decision on what tool to use for this thesis.

\section{State of the art documented}
\label{ch:tooling-literature}
The literature used several criteria for comparing mutation testing tools for Java:
\begin{itemize}
  \item Execution time of each tool.
  \item The cost as the number of test cases that needs to be generated and the the number of equivalent mutants that would have to be inspected.
  \item Effectiveness of the mutation adequate test suite of each tool. Every test of the suite adds to the effectiveness score. If a test is removed less than 100\% effectiveness can be achieved. To evaluate the effectiveness of each tool's mutation adequate test suite, a cross-testing technique is applied. The adequate test suite per tool is run on the set of mutants generated by the other tools.
\end{itemize}
Kintis et al. analysed and compared the performance of PIT,  MuJava and Major\cite{Kintis2016AnalysingStudy}. 
They started out with performing a mini literature survey on mutation testing tools used in research on test effectiveness in Java. 
With a data set of twelve methods and six Java projects they evaluated each tool using a cross-testing technique.
As a result they found that the most effective mutation adequate test suite was from MuJava, followed by Major and PIT.
The score of the application cost was the inverse of the previous ranking: PIT generated the smallest set of equivalent mutants and required the least test cases.
During their research they also compared the different mutant operators the tools supported. 
An overview of mutant operators per tool was created. 
They compared the mutant operators and reported overlapping operators.
\newline
Marki and Lindstrom performed their research on the same three mutation testing tools.\cite{Marki2017MutationJava}. 
The same cross-testing technique used by Kintis et al., was applied for three small Java programs. 
These programs are popular programs used in testing literature.
They found that the mutation tools do not subsume \todo{add subsumption to word list} each other. 
According to the research a ranking of the tools would be as follows; MuJava generated the strongest mutants followed by Major and PIT.
Furthermore they found that MuJava generated significantly more equivalent mutants.
MuJava also had an execution time twice the amount of that of Major and PIT combined.
\newline
Laurent et al., introduced PIT+, a version of PIT with extended set of mutators\cite{Laurent2017AssessingPIT}.
They used the same tests suites that were generated by kintis et al., and combined these into an adequate test suite that would detect the combined set of mutants generated by PIT, MuJava and Major. 
They discovered that the set of mutants generated by PIT+ was stronger than the combined sets generated by the other three tools.
Fortunately the extended set of mutants used in PIT+ is now integrated in the PIT project\cite{pit+}.

@@

Some of the tools have been developed further or are still actively developed\cite{pit-releases,Major}.

@@

ISI & super keyword insertion                                   & ER                    & Experimental Switch             \\ \hline
ISD & super keyword deletion                                    & BI                    & Big Integer                     \\ \hline

@@

PNC & new method call with   child class type                   & N                     & Negation                        \\ \hline

@@

JTI & this keyword insertion                                    & \multicolumn{1}{l|}{} &                                 \\ \hline
JTD & this keyword deletion                                     & \multicolumn{1}{l|}{} &                                 \\ \hline
JSI & static modifier   insertion                               & \multicolumn{1}{l|}{} &                                 \\ \hline
JSD & static modifier   deletion                                & \multicolumn{1}{l|}{} &                                 \\ \hline

@@

Tools sometimes implement the same mutator differently, resulting in different mutant sets. Table \ref{tab:mutators-overlap} shows an overview over the relations between the mutators of MuJava and PIT. This overview does include the mutators that have been reviewed in the studies of Kintis et al., and Marki and Lindstrom. However some of the new mutators map to the existing mutators.

@@

\end{tabular}
\caption{\label{tab:mutators-overlap}Overview of mutator overlaps between  MuJavav and PIT}
\end{table}
Next we tried to generate mutants with all three tools. 
All of the sources that we used for generating mutants are build with the build tools Maven or Gradle. 
Marki and Lindstrom\cite{Marki2017MutationJava} mention that Major includes support for Maven, but it is not documented anywhere.
For this reason we could not generate mutants with Major.
Table \ref{tab:mutants-generated} shows the amount of mutants generated per project and per tool. The results show that in all of the cases PIT generates more mutants than MuJava. In some cases the difference is significant. This difference can partially be explained by the fact that MuJava does not support source projects with Java version 1.7 or higher\cite{mujava}. All the the source code that uses features from this version of Java and higher resulted in an error during mutant generation.

@@

\section{Mutation tool}
We have studied literature on three different tools and extended the research with more recent data.

@@

These are test engines\todo{add to vocabulary} and are crucial for executing tests.

@@

Conforming to these requirements would recude the set of projects we could use in or experiments.

@@

For example PIT provides a plugin system in which you can inject your own code in various stages of mutation testing\cite{pit}.

@@

Based on the information presented in this chapter we dcecided that PIT would be the best choice for generating and executing mutants.

>>>>

\chapter{Mutation testing tools}
\label{ch:mutation-tool}

Before we can start clustering mutants we need to select a tool that helps us generate mutants. 
Our goal is to find the tool that generates the most amount of mutants. 
\newline
First, we review existing literature. 
Secondly we extend the comparison by adding information about the current state of the tools.
With the combination of the information gathered we make a decision on what tool to use for this thesis.

\section{State of the art documented}
\label{ch:tooling-literature}
The literature used several criteria for comparing mutation testing tools for Java:
\begin{itemize}
  \item Execution time of each tool.
  \item The cost as the number of test cases that needs to be generated and the the number of equivalent mutants that would have to be inspected.
  \item Effectiveness of the mutation adequate test suite of each tool. Every test of the suite adds to the effectiveness score. If a test is removed less than 100\% effectiveness can be achieved. To evaluate the effectiveness of each tool's mutation adequate test suite, a cross-testing technique is applied. The adequate test suite per tool is run on the set of mutants generated by the other tools.
\end{itemize}
Kintis et al. analysed and compared the performance of PIT,  MuJava and Major\cite{Kintis2016AnalysingStudy}. 
They started out with performing a mini literature survey on mutation testing tools used in existing research on test effectiveness in Java. 
With a data set of twelve methods and six Java projects they evaluated each tool using a cross-testing technique.
As a result they found that the most effective mutation adequate test suite was from MuJava, followed by Major and PIT.
The score of the application cost was the inverse of the previous ranking: PIT generated the smallest set of equivalent mutants and required the least test cases.
During their research they also compared the different mutant operators the tools supported. 
An overview of mutant operators(also known as mutators) per tool was created. 
They compared the mutant operators and reported overlapping operators.
\newline
Marki and Lindstrom performed their research on the same three mutation testing tools.\cite{Marki2017MutationJava}. 
The same cross-testing technique used by Kintis et al., was applied for three small Java programs. 
These programs are popular programs used in testing literature.
They found that the mutation tools do not subsume each other. 
According to the research a ranking of the tools would be as follows; MuJava generated the strongest mutants followed by Major and PIT.
Furthermore they found that MuJava generated significantly more equivalent mutants\cite{Marki2017MutationJava}.
MuJava also had an execution time twice the amount of that of Major and PIT combined.
\newline
Laurent et al., introduced PIT+, a version of PIT with extended set of mutators\cite{Laurent2017AssessingPIT}.
They used the same tests suites that were generated by kintis et al., and combined these into an adequate test suite that would detect the combined set of mutants generated by PIT, MuJava and Major. 
They discovered that the set of mutants generated by PIT+ was stronger than the combined sets generated by the other three tools.
Fortunately the extended set of mutants used in PIT+ is now integrated in the PIT project\cite{pit+}.

@@

Some of the tools were updated or are still under active development\cite{pit-releases,Major}.

@@

ISI & Super keyword insertion                                   & ER                    & Experimental Switch             \\ \hline
ISD & Super keyword deletion                                    & BI                    & Big Integer                     \\ \hline

@@

PNC & New method call with   child class type                   & N                     & Negation                        \\ \hline

@@

JTI & This keyword insertion                                    & \multicolumn{1}{l|}{} &                                 \\ \hline
JTD & This keyword deletion                                     & \multicolumn{1}{l|}{} &                                 \\ \hline
JSI & Static modifier   insertion                               & \multicolumn{1}{l|}{} &                                 \\ \hline
JSD & Static modifier   deletion                                & \multicolumn{1}{l|}{} &                                 \\ \hline

@@

Tools sometimes implement the same mutator differently, resulting in different mutant sets. Table \ref{tab:mutators-overlap} shows an overview over the relations between the mutators of MuJava and PIT. 
Some of the new mutators from the tools map to the already existing mutators of the other tools.
To get a complete overview we include the mutators that have been reviewed in the studies of Kintis et al., and Marki and Lindstrom.

@@

\end{tabular}
\caption{\label{tab:mutators-overlap}Overview of mutator overlaps between  MuJavav and PIT}
\end{table}
Next we tried to generate mutants with all three tools. 
All of the sources that we used for generating mutants are build with the build tools Maven or Gradle. 
Marki and Lindstrom\cite{Marki2017MutationJava} mention that Major includes support for Maven, but it is not documented anywhere.
For this reason we could not generate mutants with Major.
Table \ref{tab:mutants-generated} shows the amount of mutants generated per project and per tool. The results show that in all of the cases PIT generates more mutants than MuJava. In some cases the difference is significant. This difference can partially be explained by the fact that MuJava does not support source projects with Java version 1.7 or higher\cite{mujava}. All the code that uses the of features from this version of Java and above resulted in an error during mutant generation.

@@

\section{Selecting mutation testing tool}
We have studied literature on three different tools and extended the research with the current state of the art.

@@

These are test engines and are crucial for executing tests.

@@

Conforming to these requirements would reduce the set of projects we could use in or experiments.

@@

For example PIT provides a plugin system in which you can inject your own code in various stages of mutation testing process\cite{pit}.

@@

Based on the information presented in this chapter we decided that PIT would be the best choice for generating and executing mutants.
